---
title: "Model Card: predicting employee attrition"
date: last-modified
execute:
  echo: false
format: html
---

```{r}
#| include: false
library(tidymodels)
library(vetiver)
library(pins)
theme_set(theme_light())

b <- board_connect()
v <- vetiver_pin_read(b, "julia.silge/attrition-rf")
v_meta <- pin_meta(b, "julia.silge/attrition-rf")
```

A [model card](https://doi.org/10.1145/3287560.3287596) provides brief, transparent, responsible reporting for a trained machine learning model.

## Model details

- Developed by Julia Silge for the Imaginary HR Department.
- `r cli::pluralize("{v$description} using {ncol(v$prototype)} feature{?s}")`.
- This model predicts how likely an employee is to leave their job at our Imaginary Company, based on their income, job satisfaction, department, and whether they work overtime.
- Version `r v$metadata$version` of this model was published at `r v_meta$created`.
- If you have questions about this model, please contact the good folks in the HR department.

## Intended use

- The primary intended use of this model is to provide a useful score for identifying employees at higher risk of attrition, for interventions targeted to groups.
- The primary intended users of this model are HR leaders, to identify what kinds of employees need interventions in their work life.
- Some use cases are out of scope for this model, such as providing these scores to individual managers to be used for compensation decisions.

## Important factors

- Specific factors important to the context of this model are the demographic distribution of our employees, and how employees of different backgrounds are distributed within the company.
- In evaluating this model, we examined aspects such as how this model performs across employees in different departments and of different demographic characteristics.

## Metrics

- The metrics used to evaluate this model are accuracy and mean log loss.
- These metrics are computed via the yardstick package from the tidymodels ecosystem.
- We chose these metrics because they help us understand what proportion of our predictions are correct and how close our predictions are to the true values.

## Training data & evaluation data

- The training dataset for this model was 881 employees across all departments in our organization.
- The training dataset for this model has the "prototype" or signature:

```{r}
glimpse(v$prototype)
```

- The evaluation dataset used in this model card is 589 of the same types of employees.
- We chose this evaluation data as a random sample heldout from the original 1470 employees.

## Quantitative analyses

```{r}
## compute predictions for your evaluation data
attrition <- nanoparquet::read_parquet("attrition-eval.parquet")
preds <- augment(v, attrition)
```


### Overall model performance

```{r}
inspect_metric_set <- metric_set(accuracy, mn_log_loss)
inspect_metric_set(preds, truth = Attrition, estimate = .pred_class, .pred_No)
```

::: {.panel-tabset}

### Model performance by gender

```{r}
preds |>
  group_by(Gender) |>
  inspect_metric_set(truth = Attrition, estimate = .pred_class, .pred_No)
```

### Model performance by education

```{r}
preds |>
  group_by(Education) |>
  inspect_metric_set(truth = Attrition, estimate = .pred_class, .pred_No)
```

:::

### Visualize model performance

```{r}
preds |>
  group_by(Department) |>
  roc_curve(Attrition, .pred_No) |>
  autoplot()
```


## Ethical considerations

- We considered how this model may exacerbate existing inequality in our company, by performing better or worse for different groups.
- We do not observe significant differences in model performance by gender or department.
- We do observe differences in model performance by education, with worse performance for less educated employees. We identify this is as the most significant ethical challenge with using this model.

## Caveats & recommendations

- This model cannot be used to account for individual variation in likelihood to leave Imaginary Company.
- We recommend that upper management keep in mind that scores from this model are not equally accurate for all employees.
- This model is less useful in predicting attrition for employees with less education than those with more education.





